---
title: 'Part 6: Over 10K - All Years'
output:
  html_document:
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 

### Background

The following is summary information on complete cases with salaries over $10,000. 

### Libraries

Load required libraries.

```{r 0-libraries, warning=FALSE, message=FALSE}
library(proto)
library(gsubfn)
library(RSQLite)
library(knitr)
library(data.table)
library(plyr)
library(dplyr)
library(sqldf)
library(ggplot2)
library(RColorBrewer)
library(grid)
library(ggrepel)
library(plotly)
library(kableExtra)
library(tidyr)

# required files chunk 1
# df10.csv

# required files chunk 12
```

# Import and clean 

### Import complete cases with salaries over 10K.

### Clean data 

Remove all cases where Community Boards are the agencies. This is to provide a more accurate list of agencies as there are 59 Community Boards and each is considered its own agency, although in reality these are small offices.

Change the value of the Administrative Staff Analyst title to remove the trailing parenthesis.

View the full dataframe. It includes 672578 observations of  22 variables.

```{r 1}
#import csv of all complete cases over 10k
df10<- data.frame(read.csv("df10.csv"),stringsAsFactors = FALSE, strip.white=TRUE)

#rename dataframe
df10a<-df10

#remove community boards
df<-df10a[- grep("COMMUNITY BOARD", df10a$agency),]

#change class to character
df$Description <- as.character(df$Description)

#clean title description to remove parenthesis
df$Description[df$Description == "ADMINISTRATIVE STAFF ANALYST ("] <- "ADMINISTRATIVE STAFF ANALYST"

# Intial results indicated this may need cleanup because there are not 84 agencies-- some spaces in the agency names are causing agencies to appear twice. The following trim function successfully reduced the number of unique agencies.
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
df$agency <- trim(df$agency)

#show structure
str(df)

```

# All titles

### Titles available in the most agencies

```{r 2}

dfallyears<-df

df %>%
  group_by(Description) %>%
  summarise(count = n_distinct(agency))%>% 
  arrange(desc(count))%>%
  top_n(20)%>%
  kable("html") %>%
  kable_styling() 



```


### All titles - highest paid titles in 2017

```{r 3, eval=TRUE}

#salaries above 10k

# highest paid titles in 2017
df17<- dfallyears[dfallyears$year==2017,]

topsal<-df17 %>%
  group_by(Description) %>%
  mutate(Count = n()) %>%
  group_by(title, Description,  Count) %>%
  summarise_at(vars(sal), funs(mean(., na.rm=TRUE)  ))%>%
  arrange(desc(sal))

kable(head(topsal,20))



```

### All titles - highest paid titles in 2017, excluding Commissioners

```{r 3a}

#remove commissioners
df17a<-df17[- grep("COMMISSIONER", df17$Description),]

topsal1<-df17a %>%
  group_by(Description) %>%
  mutate(Count = n()) %>%
  group_by(title, Description,  Count) %>%
  summarise_at(vars(sal), funs(mean(., na.rm=TRUE)  ))%>%
  arrange(desc(sal))

kable(head(topsal1,20))


```


```{r 3b, eval=FALSE}




```

# Select titles only


### Subset select titles for cross-agency comparison

Subset the dataframe to include only the top 20 titles represented by the most agencies. This reduces the dataframe to 116753 observations.

These 20 titles represent 116753 of the 672578 jobs from years 2013 through 2016. 

Issue: A title may have more than one title code, partly due to defunct title codes or special agency designations. To account for this, the grouping is performed on the Description value.


```{r 4}

dfsub1<- df[df$Description %in% c(
  "COMMUNITY ASSOCIATE",
  "ADMINISTRATIVE STAFF ANALYST",
  "COMMUNITY COORDINATOR",
  "PRINCIPAL ADMINISTRATIVE ASSOC",
  "CLERICAL ASSOCIATE",
  "COMPUTER SYSTEMS MANAGER",
  "EXECUTIVE AGENCY COUNSEL",
  "ADM MANAGER-NON-MGRL FRM M1/M2",
  "ASSOCIATE STAFF ANALYST",
  "COMMUNITY ASSISTANT",
  "SECRETARY",
  "ADMINISTRATIVE PUBLIC INFORMAT",
  "AGENCY ATTORNEY",
  "COMPUTER SPECIALIST (SOFTWARE)",
  "STAFF ANALYST",
  "ADMINISTRATIVE MANAGER",
  "COMPUTER ASSOC (SOFTWARE)",
  "PROCUREMENT ANALYST",
  "COMPUTER ASSOC (OPERATIONS)",
  "ACCOUNTANT"
  ), ]

str(dfsub1)


dfsub1 %>%
  group_by(Description) %>%
  summarise(count = n_distinct(agency))%>% 
  arrange(desc(count))%>%
  top_n(20)%>%
  kable("html") %>%
  kable_styling() 
```


### Select titles - highest average salaries

Group select titles with average salary and count of title.

Arrange by descending average salary across all years.


```{r 5}
salbytitle1<-dfsub1 %>%
  group_by(Description) %>%
  mutate(Count = n()) %>%
  group_by(Description,  Count) %>%
  summarise_at(vars(sal), funs(mean(., na.rm=TRUE)  ))%>%
  arrange(desc(sal))

kable(head(salbytitle1))

```

### Select titles - average salaries by title and year

Group select titles with average salary and count of title, by year.

Arrange by title.


```{r 6}
salbytitle2<-dfsub1 %>%
  group_by(Description, year) %>%
  mutate(Count = n()) %>%
  group_by(Description, year, Count) %>%
  summarise_at(vars(sal), funs(mean(., na.rm=TRUE)  ))%>%
  arrange(Description)

kable(salbytitle2)
```

### Select titles - average salaries by title and year, with year comparison

```{r 7}
salbytitle3<-salbytitle2%>%
  select(Description, year, sal)%>%
  spread(year, sal)  

kable(salbytitle3)

```


### Select titles - average salaries by agency and year, with count

Group agencies represented in these select titles with average salary and count of title, by year.

Arrange by title.

```{r 8}
salbyagency1<-dfsub1 %>%
  group_by(agency, year) %>%
  mutate(Count = n()) %>%
  group_by(agency, year, Count) %>%
  summarise_at(vars(sal), funs(mean(., na.rm=TRUE)  ))%>%
  arrange(agency)

kable(salbyagency1)
```

### Select titles - average salaries by agency and year, with year comparison

```{r 9}
salbyagency2<-salbyagency1%>%
  select(agency, year, sal)%>%
  spread(year, sal)  

kable(salbyagency2)

```

### Select titles - headcounts by agency and year, with year comparison

```{r 10}
headcountbyagency<-salbyagency1%>%
  select(agency, year, Count)%>%
  spread(year, Count)  

kable(headcountbyagency)
```

### Select titles - percent salary increases by title

```{r 11}
salbytitle3$rawdiff  <-(salbytitle3$'2017') - (salbytitle3$'2014')
salbytitle3$pctdiff<-salbytitle3$rawdiff / (salbytitle3$'2014')
kable(salbytitle3)
```

###  Select titles - percent headcount increases by agency

```{r 12}
#percent headcount increases by agency
headcountbyagency$rawdiff  <-(headcountbyagency$'2017') - (headcountbyagency$'2014')

headcountbyagency$pctdiff<-headcountbyagency$rawdiff / (headcountbyagency$'2014')

#str(headcountbyagency)
kable(headcountbyagency)

```

# Plots

###  Plot - Select titles - percent salary increases by agency

```{r 13-1}

dfraises<- data.frame(read.csv("bestraises-selecttitles.csv"),stringsAsFactors = FALSE, strip.white=TRUE)

#p2 <- ggplot(df12, aes(x = reorder(agency, -pctincrease), y = pctincrease)) +
         # geom_bar(stat = "identity")
#p2
         
         
dfraises$title <- factor(dfraises$title, levels = unique(dfraises$title)[order(dfraises$pctincrease, decreasing = TRUE)])


p1a <- plot_ly(
  data = dfraises, 
  x = ~title, 
  y = ~pctincrease, 
  type = 'bar', 
  hoverinfo='text',
  fill="#D7BDE2",
  text = ~paste(
    'Agency: ', title,
    '<br> Salary Percent Increase: ', pctincrease) ) %>%
  layout(title="Salary increases 2014-2017, top 10 titles, select titles", margin = list(b = 200), xaxis = list(tickangle = 90))

p1a

```


###  Plot - Select titles - percent headcount increases by agency

```{r 13-2}

dfheadcount<- data.frame(read.csv("headcountbyagency-pct-increase-top10.csv"),stringsAsFactors = FALSE, strip.white=TRUE)

#p2 <- ggplot(df12, aes(x = reorder(agency, -pctincrease), y = pctincrease)) +
         # geom_bar(stat = "identity")
#p2
         
         
dfheadcount$abbreviation <- factor(dfheadcount$abbreviation, levels = unique(dfheadcount$abbreviation)[order(dfheadcount$pctincrease, decreasing = TRUE)])


p1 <- plot_ly(
  data = dfheadcount, 
  x = ~abbreviation, 
  y = ~pctincrease, 
  type = 'bar', 
  hoverinfo='text',
  fill="#D7BDE2",
  text = ~paste(
    'Agency: ', agency,
    '<br> Headcount Percent Increase: ', pctincrease) ) %>%
  layout(title="Headcount increases 2014-2017, top 10 agencies, select titles", margin = list(b = 200), xaxis = list(tickangle = 90))

p1

```


###  Plot - All Titles - Top 10 agencies with the most jobs over $10K, 2017

```{r 14}
df3<- data.frame(read.csv("agency17high.csv"),stringsAsFactors = FALSE)

df3 %>% mutate_if(is.factor, as.character) -> df3

p2 <- df3 %>%
  plot_ly(labels = ~ADDRESS, values = ~n) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Agencies with the most jobs over $10K",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

p2

```


###  Plot - All Titles - Top 10 titles with the most jobs over $10K, 2017

```{r 15}
df4<- data.frame(read.csv("agency17high-titles.csv"),stringsAsFactors = FALSE)

df4 %>% mutate_if(is.factor, as.character) -> df4

p3 <- df4 %>%
  plot_ly(labels = ~Description, values = ~n) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Titles with the most jobs over $10K",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

p3

```


###  Plot - All Titles - Histogram: Distribution of 10K+ salaries, 2017

```{r 16}

hist(df17$sal, 
     main ="Distribution of salaries over $10K, 2017",
     xlab="Salary",
     border="black",
     col="#58D68D",
     xlim=c(0,200000),
     ylim=c(0,35000),
     breaks = seq(0, 350000, by = 5000)
     #prob=TRUE
     #breaks=c(0,10000,20000,30000,40000,50000,60000,70000,80000,90000,100000,110000,120000,130000,350000)
     )


```

###  Plot - All Titles - Salaries between 85K and 90K

What constitutes the spike in the histogram between \$85,000 and \$95,000?

```{r 17}
df5<- data.frame(read.csv("titles17-8590k.csv"),stringsAsFactors = FALSE)

df5 %>% mutate_if(is.factor, as.character) -> df5

p4 <- df5 %>%
  plot_ly(labels = ~Description, values = ~n) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Titles between $85K and $90K",  showlegend = T,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

p4

```


###   Plot - Which titles have employees across the most agencies, excluding Community Boards?

###   Plot - For select titles, do some agencies pay more?

###   Plot - Titles with the greatest salary ranges 

Best opportunities for promotion within the same title


# Correlations

###   Correlation with budget

###   Correlation with unions

# Export files

Export:

Subset of selected 20 titles.

Summary of salaries by title with year over year comparisons.

Summary of salaries by agency with year over year comparisons.

Headcount by agency, wide format

Average salary by title, long format

Headcount and average salary by title, long format

Percent headcount increases by agency

2017 payroll only, complete cases, all titles


```{r 20, eval=FALSE}
#changed to eval=FALSE after first run.

#subset of selected titles only
write.csv(dfsub1, file = "dfsub1.csv")

#average salary comparison with one column for each year
write.csv(salbytitle3, file = "sal-title-year-comparison.csv")

#headcount by agency with one column for each year
write.csv(headcountbyagency, file = "headcount-agency-year-comparison.csv")

#average salary comparison with years in one column
write.csv(salbytitle2, file = "sal-title-year-comparison-long.csv")

#headcount and average salary by agency with years in one column
write.csv(salbyagency1, file = "sal-agency-year-comparison-long.csv")

#percent headcount increases by agency
write.csv(headcountbyagency, file = "headcountbyagency-pct-increase.csv")

#2017 payroll only, complete cases, all titles
write.csv(df17, file = "df17.csv")

```
