---
title: 'Part 4: Analysis'
output:
  html_document:
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries

```{r 0-libraries, warning=FALSE, message=FALSE}
library(proto)
library(gsubfn)
library(RSQLite)
library(knitr)
library(data.table)
library(plyr)
library(dplyr)
library(sqldf)
library(ggplot2)
library(RColorBrewer)
library(grid)
library(ggrepel)
library(plotly)
library(kableExtra)
```


# Table: Which titles have employees across the most agencies, excluding agencies which are Community Boards?

These could be considered the most flexible titles, for employees who may want to move around across agencies while retaining their civil service title.

```{r 8a}
df15<- data.frame(read.csv("df15.csv"),stringsAsFactors = FALSE)
df7<-df15

df7a<-df7[- grep("COMMUNITY BOARD", df7$ADDRESS),]

df7a %>%
  group_by(Description) %>%
  summarise(count = n_distinct(ADDRESS))%>% 
  arrange(desc(count))%>%
  top_n(20)%>%
  kable("html") %>%
  kable_styling() 
```

# Bar Plot: Which titles have employees across the most agencies, excluding agencies which are Community Boards?

```{r 8b, eval=FALSE}
# This keeps all 1089 descriptions from the original data frame
df7b<-df7a %>%
  group_by(Description) %>%
  summarise(count = n_distinct(ADDRESS))%>% 
  arrange(desc(count))%>%
  top_n(10)%>%
  as.data.frame()

str(df7b)

p7 <- plot_ly(
  data = df7b, 
  x = ~Description, 
  y = ~count, 
  type = 'bar', 
  #mode = 'markers',
  hoverinfo='text',
  text = ~paste(
    'Title: ', Description,
    '<br> Employees: ', count) ) %>%
  layout(title="Titles across the most agencies, excluding Community Boards")

p7
```

```{r 8c}

df7b<- data.frame(read.csv("titlesmultipleagencies.csv"),stringsAsFactors = FALSE)

df7b$Description <- factor(df7b$Description, levels = unique(df7b$Description)[order(df7b$COUNT, decreasing = TRUE)])


p7 <- plot_ly(
  data = df7b, 
  x = ~Description, 
  y = ~COUNT, 
  type = 'bar', 
  hoverinfo='text',
  fill="#D7BDE2",
  text = ~paste(
    'Title: ', Description,
    '<br> Agencies: ', COUNT) ) %>%
  layout(title="Titles across the most agencies, excluding Community Boards", margin = list(b = 200), xaxis = list(tickangle = 90))

p7

```


# Table: What are the highest average salaries by agency and title?

```{r 9, eval=FALSE}

summary(df7a)

df7a %>%
  group_by(Description) %>%
  summarise(count = n_distinct(ADDRESS), avgsal=mean(SAL))%>% 
  arrange(desc(avgsal))%>%
  top_n(20)%>%
  kable("html") %>%
  kable_styling() 
```

```{r 9a, eval=TRUE}

df7c <- filter(df7a, 
                 TITLE == "56057" | 
                 TITLE == "10124" | 
                 TITLE == "10251" | 
                 TITLE == "56058" | 
                 TITLE == "10026" | TITLE == "1002A" | 
                 TITLE == "10050" | 
                 TITLE == "95005" | 
                 TITLE == "12627" | 
                 TITLE == "10252" | 
                 TITLE == "1002C"  
                 )

# 56057 COMMUNITY ASSOCIATE
# 10124 PRINCIPAL ADMINISTRATIVE ASSOC
# 10251 CLERICAL ASSOCIATE
# 56058 COMMUNITY COORDINATOR
# 10026, 1002A ADMINISTRATIVE STAFF ANALYST (
# 10050 COMPUTER SYSTEMS MANAGER
# 95005 EXECUTIVE AGENCY COUNSEL
# 12627 ASSOCIATE STAFF ANALYST
# 10252 SECRETARY
# 1002C ADM MANAGER-NON-MGRL FRM M1/M2


df7c %>%
  group_by(ADDRESS, Description) %>%
  summarise(count = n(), avgsal=mean(SAL))%>% 
  arrange(Description, desc(avgsal))%>%
  top_n(20)%>%
  kable("html") %>%
  kable_styling()  

df7csv<-df7c %>%
  group_by(ADDRESS, Description) %>%
  summarise(count = n(), avgsal=mean(SAL))%>% 
  arrange(Description, desc(avgsal))

#write.csv(df7csv,'df7csv.csv')
```


# For titles across multiple agencies, do some agencies hire the same title at a higher average salary?

```{r 9b}

df7d<- data.frame(read.csv("topsalagencies-bytitles.csv"),stringsAsFactors = FALSE)

df7d$Agency <- factor(df7d$Agency, levels = unique(df7d$Agency)[order(df7d$COUNT, decreasing = TRUE)])


p7d <- plot_ly(
  data = df7d, 
  x = ~Agency, 
  y = ~COUNT, 
  type = 'bar', 
  hoverinfo='text',
  fill="#76D7C4",
  text = ~paste(
    'Agency: ', Agency,
    '<br> Times Appearing in Top 10: ', COUNT) ) %>%
  layout(title="Highest Paying Agencies for Select Titles, Excluding Community Boards", margin = list(b = 200), xaxis = list(tickangle = 90, title=""))

p7d

```


# What titles have the most opportunities for salary increases?
## Min and max in each title

# What are the most significant changes between 2014 and 2015?

# Is there a correlation between union representation and salary?

# Can this data predict current job postings in 2018? 

# Is there a relationship between an agency's budget and its average salaries?